Historical Soil Organic Carbon Budget

### 2020-01-28 SOCbudgetPaper/SOCbudget.Rmd
Rewrite of methods part 1 on Calculating carbon stocks

  Please add `\clearpage` between each table and/or figure. Further guidelines on figures and tables can be found below.
  Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
  To rename them correctly to A1, A2, etc., please add the following commands in front of them:
output:
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
  rticles::copernicus_article: default
  md_document:
    variant: markdown_github
---

```{r child = 'introduction.Rmd'}
```

```{r child = 'methods.Rmd'}
```

```{r child = 'results.Rmd'}
```

```{r child = 'discussion.Rmd'}
```

```{r child = 'conclusion.Rmd'}
```

>>>>

  Please add `\clearpage` between each table and/or figure. Further guidelines on figures and tables can be found below.
  Regarding figures and tables in appendices, the following two options are possible depending on your general handling of figures and tables in the manuscript environment:
  To rename them correctly to A1, A2, etc., please add the following commands in front of them:
output:
  bookdown::pdf_book:
    base_format: rticles::copernicus_article # for using bookdown features like \@ref()
    md_extensions: +raw_attribute
  rticles::copernicus_article: default
  md_document:
    variant: markdown_github
---
\newpage

```{r child = 'introduction.Rmd'}
```
\newpage

```{r child = 'methods.Rmd'}
```
\newpage

```{r child = 'results.Rmd'}
```
\newpage

```{r child = 'discussion.Rmd'}
```
\newpage

```{r child = 'conclusion.Rmd'}
```
\newpage

### 2020-01-28 SOCbudgetPaper/SOCbudget.tex
Rewrite of methods part 1 on Calculating carbon stocks

\correspondence{Kristine\ Karstens\ (kristine.karstenst@pik-potsdam.de)}

@@

@@

Introduction text goes here. You can change the name of the section if
neccessary using
\texttt{\textbackslash{}introduction{[}modified\ heading{]}}.

\section{Method (50)}

\subsection{Carbon Budget (50)}

Based on the IPCC guidelines vol.~4 and their refinement
(\citep{ipcc_2006_2006},\citep{ipcc_2019_2019}) for soil organic carbon
stocks, we combining approaches to estimate SOC stocks by weighting
inflows via dead plant material (see \ref{sec:carboninputs}) against
outflows through SOC decay (see \ref{sec:tier2}). Carbon displacement
via leaching and erosion is neglected in this study. We calculate annual
land use type specific budgets for cropland, pastures and natural
vegetation, also representin land conversion as transfer between landuse
type budgets. A simple approach based o the tier 1 method of the older
IPCC guidelines vol 4. (\citet{ipcc_2006_2006}) using stock change
factors, is applied to cross validated results \ref{sec:tier1}

\subsubsection{Carbon Inputs to the Soil}\label{sec:carboninputs}

Carbon input estimations are based on the land use type. Whereas
cropland inputs are mainly formed by disaggregated country statistics on
residue, dead below ground and cover crop biomass, pasture and natural
vegetation inputs are estimate via modelled annual litterfall rates.
Using the steady-state method of the IPCC guidelines
(\citep{ipcc_2019_2019}) carbon inputs have to be accompined by data on
lignin and nitrogen content to allocate dead plant biomass to the
corrosponding soil pools based on the chemical texture. Sources for all
use data can be found in table \ref{tab:datasourceinputs}

\begin{table*}[h]
\caption{Sources for carbon input data}
\begin{tabular}{l l l}
\tophline
 Land use types   & carbon inputs & nitrogen and lignin content \\
\middlehline
 Cropland         & FAO statistics, AQUASTAT, LPJmL4 [1] & default values [\cite{ipcc_2006_2006}] \\
 Pasture          & annual litterfall in $\tfrac{gC}{m^2}$ from LPJmL4 - manage grassland [3] & default values [2] \\
 Natural vegetation & annual litterfall in $\tfrac{gC}{m^2}$ from LPJmL4  & \begin{minipage}[t]{0.37\columnwidth}\raggedright\strut Nitrogen and lignin content of tree compartments used in CENTURY \strut \end{minipage}\tabularnewline
\bottomhline
\end{tabular}
\label{tab:datasourceinputs}
\belowtable{}
\end{table*}

\subsubsection{Soil Carbon turnover following (new) Tier 2 method
(300)}\label{sec:tier2}

We are following the steady-state method of the refinement of the IPCC
guidelines vol.~4 (\citet{ipcc_2019_2019}) by calculating yearly
turnover and transfer rates between three different SOC pools for the
topsoil (0-30 cm). The approach is based on global parameters
({[}@ref(ipcc\_2019\_2019){]}) as well as half-degree data on sand
fraction (SoilGrids), temperature, preciptation and potential
evapotranspiration (CRU). Next to the given climatic and natural
biophysical properties irrigation regime (rainfed vs.~irrigated) as well
as tillage (as soil disturbance indicator) modificate processes. For
cropland an assessment of tillage types and irrigation conditions has
been made, whereas on pastures and natural vegetation, we assume rainfed
and non-tilled conditions.

\subsubsection{Soil Carbon turnover following Tier 1
(150)}\label{sec:tier1}

@@

\subsection{Agricultural management (50)}\label{sec:agrimanagement}

@@

\subsubsection{Landuse and Landuse Change (150)}

We use LUH2v2 data for major Landuse types and their transition and fit
cropspecific areas to country scale FAO data.

\subsubsection{Crop Production and Residues (300)}

FAO Production values are combined with Feed estimations from
{[}Isabelles Paper{]} and rule based demand shares. LPJmL yield and LUH
landuse patterns are used to scale down to half-degree.

\subsubsection{Livestock Distribution and Manure Excretion (300)}

@@

\subsubsection{Irrigation (100)}

@@

\subsubsection{Tillage (100)}

@@

\section{Results}

\section{Discussion}

@@

The conclusion goes here. You can modify the section name with
\texttt{\textbackslash{}conclusions{[}modified\ heading\ if\ necessary{]}}.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Figures and tables in appendices}\subsection{Option 1}

If you sorted all figures and tables into the sections of the text,
please also sort the appendix figures and appendix tables into the
respective appendix sections. They will be correctly named
automatically. \subsection{Option 2} If you put all figures after the
reference list, please insert appendix tables and figures after the
normal tables and figures.

>>>>

\correspondence{Kristine\ Karstens\ (\href{mailto:kristine.karstenst@pik-potsdam.de}{\nolinkurl{kristine.karstenst@pik-potsdam.de}})}

@@

\newpage

@@

Introduction text goes here.
You can change the name of the section if neccessary using \texttt{\textbackslash{}introduction{[}modified\ heading{]}}.
\newpage

\hypertarget{method-50}{%
\section{Method (50)}\label{method-50}}

\hypertarget{sec:carbonbudget}{%
\subsection{Carbon Stocks following (new) Tier 2 method (50)}\label{sec:carbonbudget}}

We calculate annual land use type specific soil organic carbon stocks for cropland, pastures and natural vegetation on half-degree resolution for the period of 1965 to 2010 based on the following three steps: (1) Calculating the land use (sub-)type specific steady-states and decay rates for SOC stocks given the current biophysical, climatic and agronomic conditions, (2) accounting for land conversation effects by transferring SOC between land use types and (3) updating SOC stocks based on the previous stock, the steady-state and the decay rate.

\hypertarget{steady-state-soc-stocks-and-decay-rates}{%
\subsubsection{Steady-state SOC stocks and decay rates}\label{steady-state-soc-stocks-and-decay-rates}}

In a simple first order kinetic approach the steady-state soil organic carbon stocks \(SOC^{eq}\) are given by
\begin{equation}
SOC^{eq} =\frac{C^{\textrm{in}}}{k}
\label{eq:inoutflow}
\end{equation}
with \(C^{\textrm{in}}\) being carbon inputs to the soil and \(k\) denoting the soil organic carbon decay rate. We use for our calculations the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils, which assume three soil carbon pools and entangled dynamics between them. Carbon inflow to each subpool (see \ref{sec:carboninputs}) and decay rates (see \ref{sec:tier2}) of each subpool are still the key components to determining steady-state SOC stocks.

\hypertarget{sec:carboninputs}{%
\subsubsection{Carbon Inputs to the Soil}\label{sec:carboninputs}}

We account for different carbon inputs sources depending on the land use type (see table \ref{tab:datasourceinputs}). Following the IPCC methodology carbon inputs are disaggregated into different structural components depending on their lignin and nitrogen content (see @ref(ipcc\_2019\_2019)). For each structural components the sum over all carbon inputs sources is allocated to the respective SOC sub pools. This implies that not only the amount of carbon, but also their structural composition is determining the effective inflow. Data sources for all considered carbon inputs as well as for lignin and nitrogen content can be found in table \ref{tab:datasourceinputs}.

 \begin{table*}[h]
 \caption{Type and data sources for carbon inputs to different land use types}
 \begin{tabular}{l l l l}
 \tophline
  Land use types   & source of carbon inputs & data source & nitrogen and lignin content \\
 \middlehline
 \multirow{3}{*}{Cropland} & residues & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2]  \\
                            & dead below ground biomass of crops & FAOSTAT, LPJmL4 [2, sec:residues] & default values given by [2] \\
                            & manure & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
                            \hline
 \multirow{2}{*}{Pasture}  & annual litterfall & LPJmL4 [3] & default values given by [2] \\ 
                            & manure  & FAOSTAT, Isabelle [2, sec:manure] & default values given by [2] \\
                            \hline
  Natural vegetation        & annual litterfall & LPJmL4 [4]& \begin{minipage}[t]{0.28\columnwidth}\raggedright\strut Nitrogen and lignin content of tree compartments used in CENTURY [4] \strut \end{minipage}\tabularnewline
 \bottomhline
 \end{tabular}
 \label{tab:datasourceinputs}
 \belowtable{}
 \end{table*}

\hypertarget{sec:tier2}{%
\subsubsection{Soil Organic Carbon decay (300)}\label{sec:tier2}}

Decay rates are influenced by climatic conditions, biophysical and biochemical soil properties as well as management factors that vary over time (t) and space (i). Following the steady-state method of the refinement of the IPCC guidelines vol.~4 (\citet{ipcc_2019_2019}) for mineral soils we consider temperature (temp), water (wat), sand fraction (sf) and tillage (till) effects to spatially disaggregate default global decay rates \(k_{sub}\) given by the IPCC via

\begin{equation}
k_{sub,t,i} = \quad k_{sub} \quad \cdot temp_{t,i} \quad \cdot wat_{t,i} \underbrace{\cdot sf_{t,i} \quad \cdot till_{t,i}.}_{\text{only included for some subpools}}
\label{eq:decayrates}
\end{equation}

For cropland we performed an assessment of tillage types and irrigation conditions, whereas on pastures and natural vegetation, we assume rainfed and non-tilled conditions. Data sources as well as considered effects for each land use types are shown in table \ref{tab:datasourcedecay}. To account for variations of decay rates within grid cells based on different tillage and irrigation regimes, average rates based on area shares are calculated.

 \begin{table*}[h]
 \caption{Type and data sources for carbon inputs to different land use types}
 \begin{tabular}{l l l l}
 \tophline
  Land use types   & type of decay driver & parameter use to represent driver & data source \\
 \middlehline
 \multirow{2}{*}{all} & Soil quality & Sand fraction of the first 0-30 cm &  [SoilGrids]  \\
                      \cline{2-4}

                      & Mircobial activity & air temperature & [CRUp4.0] \\
                      \cline{2-4}

                      & Water restriction & precipitation \& potential evapotranspiration & [CRUp4.0] \\
                      \cline{1-4}
\multirow{2}{*}{\begin{minipage}[t]{0.2\columnwidth}\raggedright\strut Cropland\\(additionally)\strut\end{minipage}} & Water restriction*  & irrigation  & [sec:irrigation] \\ 
                      \cline{2-4}

                      & Soil disturbance & tillage & [sec:tillage] \\
 \bottomhline
 \end{tabular}
 \label{tab:datasourcedecay}
 \belowtable{}
 \end{table*}

\hypertarget{soc-transfer-between-land-use-types}{%
\subsubsection{SOC transfer between land use types}\label{soc-transfer-between-land-use-types}}

We calculate SOC stocks based on the area shares of land use types (lut) within the half-degree grid cells (i). If land is converted from one land use type into another, a respective share of the SOC stocks have to be reallocated as well. We account for land conversion at the beginning of each time step \(t\) by calculating a preliminary stock \(SOC_{lut,t*}\) via

\begin{equation}
SOC_{lut,t*} = SOC_{lut,t-1} - \frac{SOC_{lut,t-1}}{A_{lut,t-1}} \cdot  AR_{lut,t} + \frac{SOC_{!lut,t-1}}{A_{!lut,t-1}} \cdot  AE_{lut,t} \qquad \forall sub, i  
\label{eq:ctransfer}
\end{equation}

with \(A\) being the area, \(AR\) the area reduction and \(AE\) the area expansion for a given land use type \(lut\). Note that \(!lut\) denotes the (sum over) land use type(s), which decrease(s) in the specific time step \(t\). Data sources and methodology on land use states and changes are descripted in \ref{sec:landuse}.

\hypertarget{total-soc-stocks}{%
\subsubsection{Total SOC stocks}\label{total-soc-stocks}}

Carbon stocks \(SOC\) for each subpool (sub) converge towards the calculate steady-state stock \(SOC^{eq}\) for each land-use types (lut), each subpool (sub) and each annual time step (t) as represented in equation \eqref{eq:steadystate}.

\begin{equation}
SOC_{t} = SOC_{t-1} + (SOC^{eq}_{t} - SOC_{t-1}) \cdot k_{t} \cdot \Delta t \qquad \forall lut, sub, i.
\label{eq:steadystate}
\end{equation}

The global SOC stock for each time step can than be calculated via

\begin{equation}
SOC_{t} = \sum_{i} \underbrace{\sum_{lut} \overbrace{\sum_{sub} SOC_{lut, sub, t, i}.}^{\text{land use type specific SOC stock within a grid cell}}}_{\text{total SOC stock within a grid cell}}
\label{eq:totalstock}
\end{equation}

\newpage

\hypertarget{sec:tier1}{%
\subsection{Carbon Budget following Tier 1 (150)}\label{sec:tier1}}

@@

\hypertarget{sec:agrimanagement}{%
\subsection{Agricultural management (50)}\label{sec:agrimanagement}}

@@

\hypertarget{sec:landuse}{%
\subsubsection{Landuse and Landuse Change (150)}\label{sec:landuse}}

We use the Land-Use Harmonization 2 (LUH2, \citep{LUH2}) half-degree data for major land-use states. Cropland is subdivided using country scale FAO data (\citep{FAOSTAT}) as well as crop type specific half-degree data from LUH2v2 to disaggregate into our 19 crop types on half degree resolution (see appendix for table on crop types and their mapping to FAOSTAT and LUH types as well as for technical discription of the disaggregation method).
Landuse transitions are calculate as area differences, allowing any land use type to either expand or shrink in each time step. Within cropland we do not track area transitions, following the averaging approach as descripted in \ref{sec:tier2}/\ref{sec:carbonbudget}

\hypertarget{crop-production-and-residues-300}{%
\subsubsection{Crop Production and Residues (300)}\label{crop-production-and-residues-300}}

FAOSTAT production (\citep{FAOSTAT}) values are combined with feed estimations including residue intake (\citep{weindl}) and rule based demand shares. LPJmL yield and calculate land use pattern (see @ref(\#sec:landuse)) are used to scale down to half-degree.

\hypertarget{livestock-distribution-and-manure-excretion-300}{%
\subsubsection{Livestock Distribution and Manure Excretion (300)}\label{livestock-distribution-and-manure-excretion-300}}

@@

\hypertarget{irrigation-100}{%
\subsubsection{Irrigation (100)}\label{irrigation-100}}

@@

\hypertarget{tillage-100}{%
\subsubsection{Tillage (100)}\label{tillage-100}}

@@

\newpage

\hypertarget{results}{%
\section{Results}\label{results}}

\newpage

\hypertarget{discussion}{%
\section{Discussion}\label{discussion}}

Shortcommings:

\begin{itemize}
\item
  Carbon displacement via leaching and erosion is neglected in this study.
\item
  Non-net/Gross land use transitions are not tracked in this study.
  \newpage
\end{itemize}

@@

The conclusion goes here.
You can modify the section name with \texttt{\textbackslash{}conclusions{[}modified\ heading\ if\ necessary{]}}.
\newpage

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\appendix
\section{Figures and tables in appendices}
\subsection{Option 1}

If you sorted all figures and tables into the sections of the text, please also sort the appendix figures and appendix tables into the respective appendix sections.
They will be correctly named automatically.

\subsection{Option 2}

If you put all figures after the reference list, please insert appendix tables and figures after the normal tables and figures.

### 2020-01-28 SOCbudgetPaper/discussion.Rmd
Rewrite of methods part 1 on Calculating carbon stocks

# Examples from the official template

## FIGURES

When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage between bibliography and first table and/or figure as well as between each table and/or figure.

## FIGURES

When figures and tables are placed at the end of the MS (article in one-column style), please add \clearpage between bibliography and first table and/or figure as well as between each table and/or figure.

### ONE-COLUMN FIGURES

Include a 12cm width figure of Nikolaus Copernicus from [Wikipedia](https://en.wikipedia.org/wiki/File:Nikolaus_Kopernikus.jpg) with caption using R Markdown.

### TWO-COLUMN FIGURES

You can also include a larger figure.

## TABLES

You can ad \LaTeX table in an R Markdown document to meet the template requirements.

### ONE-COLUMN TABLE

\begin{table}[t]
\caption{TEXT}
\begin{tabular}{l c r}
\tophline

a & b & c \\
\middlehline
1 & 2 & 3 \\

\bottomhline
\end{tabular}
\belowtable{Table Footnotes}
\end{table}

### TWO-COLUMN TABLE

\begin{table*}[t]
\caption{TEXT}
\begin{tabular}{l c r}
\tophline

a & b & c \\
\middlehline
1 & 2 & 3 \\

\bottomhline
\end{tabular}
\belowtable{Table footnotes}
\end{table*}

## MATHEMATICAL EXPRESSIONS

All papers typeset by Copernicus Publications follow the math typesetting regulations given by the IUPAC Green Book (IUPAC: Quantities, Units and Symbols in Physical Chemistry, 2nd Edn., Blackwell Science, available at: http://old.iupac.org/publications/books/gbook/green_book_2ed.pdf, 1993).

Physical quantities/variables are typeset in italic font (t for time, T for Temperature)

Indices which are not defined are typeset in italic font (x, y, z, a, b, c)

Items/objects which are defined are typeset in roman font (Car A, Car B)

Descriptions/specifications which are defined by itself are typeset in roman font (abs, rel, ref, tot, net, ice)

Abbreviations from 2 letters are typeset in roman font (RH, LAI)

Vectors are identified in bold italic font using \vec{x}

Matrices are identified in bold roman font

Multiplication signs are typeset using the LaTeX commands `\times` (for vector products, grids, and exponential notations) or `\cdot`

The character * should not be applied as mutliplication sign

## ALGORITHM

  If you want to use algorithms, you can either enable the required packages in the header (the default, see `algorithms: true`), or make sure yourself that the \LaTeX packages `algorithms` and `algorithmicx` are installed so that `algorithm.sty` respectively `algorithmic.sty` can be loaded by the Copernicus template.
Copernicus staff will remove all undesirable packages from your LaTeX source code, so please stick to using the header option, which only adds the two acceptable packages.

\begin{algorithm}
\caption{Algorithm Caption}
\label{a1}
\begin{algorithmic}
\STATE $i\gets 10$
  \IF {$i\geq 5$} 
\STATE $i\gets i-1$
  \ELSE
\IF {$i\leq 3$}
\STATE $i\gets i+2$
  \ENDIF
\ENDIF
\end{algorithmic}
\end{algorithm}

>>>>

# Discussion

Shortcommings:

- Carbon displacement via leaching and erosion is neglected in this study.

- Non-net/Gross land use transitions are not tracked in this study.
